---
title: "Lab 6"
author: "Samantha Kuglen"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Example of a rank-based test

We'll make out own samples, using a pseudorandom generator. Set seed allows us to get the starting point of getting "predictable, pseudo-random" random values. Setting the starting point of values from which they are pulled. You must clear the workspace to reset the seed
```{r}
set.seed(1414)

#20 is the high value
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)

```

```{r}
# quick historgram for exploratory vectors...
# not from normally distributed populations
hist(gp_1)

hist(gp_2)
```

Try a T-test:
```{r}
t.test(gp_1, gp_2)
```
The p-value is 0.1982. If these samples were drawn from populations with the same mean, the probability of taking two random samples with means *at least as different* as the sample means we would by random chance (taking into account spread, n) is 19.8%. There is a 19.8% chance we would have found samples as different as ours by random chance.

Retain (fail to reject) the null hypothesis >> There is no significant different in means between group 1 and group 2. Note: never say "accept" the null

Now let's compare outcome to a rank-based test.

## Mann Whitney U unpaired rank-based test
```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu

```
The p value of 0.28 means that if these samples were drawn from populations of the same ranks (medians), the probability of finding two samples with ranks *at least as different* as those in our samples is 0.28. (28% of the time). Not so unlikely to happen... so we do not reject the null hypothesis.

There is no significant difference in ranks (medians) between group 1 and group 2 (statistical summary).

Median scores for group 1 (M=14) and group 2 (M = 12) did not differ significantly (U(df) = 86, p =0.28). (Mann Whitney U test: U(df) = 86, p = 0.28)

On assignments make sure tests align with what is presented to audience, decide which metric to compare things for a consistent message to compare groups throughout analysis.



